---
title: "Principal component method"
author: "Valery Andruynichev"
date: '25 dec 2018 year '
output: html_document
---
# Comparison of the effectiveness of the method of principal components for the portfolio of the currency pairs and a separate currency pairs.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'D:/GH/Currency')
```

## Specify the working directory.
```{r, }
dir <- 'D:/GH/Currency' 
setwd(dir) 
```

### And connect the pakeges.
```{r, message=FALSE, warning=FALSE}
library(ggplot2) # Plots
library(dplyr) # data manipulation
library(ggfortify) # Correct data
library(readr) # load data
library(xts) # work with .xts
library(formattable) # Correct data
library(tseries) # Statistical test
```

### Load necessary data.
```{r}
# USD/RUB data
U_R <- read_delim("USDRUB_TOM.csv",";", escape_double = FALSE, col_types = cols(`AV ETC` = col_character(),Close = col_character(), Date = col_date(format = "%d.%m.%Y"),Deal = col_character(),Max = col_character(),Min = col_character(), Open = col_character(),Volume = col_character()),trim_ws = TRUE)
# EUR/RUB data
E_R <- read_delim("EURRUB_TOM.csv",";", escape_double = FALSE, col_types = cols(`AV ETC` = col_character(),Close = col_character(), Date = col_date(format = "%d.%m.%Y"),Deal = col_character(),Max = col_character(),Min = col_character(), Open = col_character(),Volume = col_character()),trim_ws = TRUE)
# GBP/RUB data
G_R <- read_delim("GBPRUB_TOM.csv",";", escape_double = FALSE, col_types = cols(`AV ETC` = col_character(),Close = col_character(), Date = col_date(format = "%d.%m.%Y"),Deal = col_character(),Max = col_character(),Min = col_character(), Open = col_character(),Volume = col_character()),trim_ws = TRUE)
# CNY/RUB data
Y_R <- read_delim("CNYRUB_TOM.csv",";", escape_double = FALSE, col_types = cols(`AV ETC` = col_character(),Close = col_character(), Date = col_date(format = "%d.%m.%Y"),Deal = col_character(),Max = col_character(),Min = col_character(), Open = col_character(),Volume = col_character()),trim_ws = TRUE)
```

### Combine the data into a single table, take the closing prices.
```{r message=FALSE, warning=FALSE}
# Leave the closing price
U_R <- U_R[,c(1,6)];E_R <- E_R[,c(1,6)];G_R <- G_R[,c(1,6)];Y_R <- Y_R[,c(1,6)]
# Combine data
df <- merge(x = U_R, E_R, by = 'Date')
df <- merge(x = df, G_R, by = 'Date')
df <- merge(x = df, Y_R, by = 'Date')
colnames(df) <- c("Date","USD/RUB","EUR/RUB","GBT/RUB","CNY/RUB")
head(df)
```

### Check the data for correctness.
```{r}
# Check the column classes
sapply(df, function(x) {class(x)})
# Check for NA values
sapply(df, function(x) {sum(is.na(x))})
```

### There are no missing (NA) values in the sample. But the data is in the wrong format.
### To convert the data to the required format, replace commas "," with dots ".".
```{r}
df$`USD/RUB` <- gsub(',','.',df$`USD/RUB`)
df$`EUR/RUB` <- gsub(',','.',df$`EUR/RUB`)
df$`GBT/RUB` <- gsub(',','.',df$`GBT/RUB`)
df$`CNY/RUB` <- gsub(',','.',df$`CNY/RUB`)
head(df)
```

### Change the data class to numeric.
```{r}
df$`USD/RUB` <- as.numeric(df$`USD/RUB`)
df$`EUR/RUB` <- as.numeric(df$`EUR/RUB`)
df$`GBT/RUB` <- as.numeric(df$`GBT/RUB`)
df$`CNY/RUB` <- as.numeric(df$`CNY/RUB`)
sapply(df, function(x) {class(x)})
```
The data corresponds to the required class.

### Consider the correct data in tabular and graphical form.
```{r}
# Replace zero values with previous ones
for (i in 2:nrow(df)) {  
  df[i,2] = ifelse(df[i,2] == 0, df[c(i-1),2], df[i,2]) 
  df[i,3] = ifelse(df[i,3] == 0, df[c(i-1),3], df[i,3]) 
  df[i,4] = ifelse(df[i,4] == 0, df[c(i-1),4], df[i,4]) 
  df[i,5] = ifelse(df[i,5] == 0, df[c(i-1),5], df[i,5]) 
}

plot(df$`USD/RUB`, type = 'l', ylab = 'USD/RUB')
plot(df$`EUR/RUB`, type = 'l', ylab = 'EUR/RUB')
plot(df$`GBT/RUB`, type = 'l', ylab = 'GBT/RUB')
plot(df$`CNY/RUB`, type = 'l', ylab = 'CNY/RUB')
```

## Let's analyze the data using the principal components method for a portfolio consisting of currency pairs.

### Find the main components for currency pairs.
```{r}
df_log <- df[,-1] # Remove the year data
pc <- prcomp(df_log, scale. = TRUE) # Finding the main components
```
Average values for currency pairs.
```{r}
pc$center
```
The value of the component for currency pairs.
```{r}
pc$scale
```
Consider the component.
```{r}
head(pc$x)
```
Let's define the variance values for each component.
```{r}
v <- summary(pc)$importance
v
```
The table shows that the main contribution to the change in the value of the asset brings component PC1 `r formattable::percent(v[3])`, component PC2 vnosti only `r formattable::percent(v[3,2])` variance. Together they explain 96.6% of the variance. Therefore, these are the components we can call the main, there is no need to consider more than two components.

### Reflect the percentage of variance explained for each component in the graph.
```{r}
pc
pc_var <- pc$sdev^2
prop_var <- pc_var/sum(pc_var)
plot(prop_var, xlab = "PC", ylab = "Proportion of variance explained", type = "b")
plot(cumsum(prop_var), xlab = "PC",
     ylab = "The cumulative proportion of explained variation",
     type = "b")
autoplot(pc, loadings = TRUE, loadings.label = TRUE)
```

According to the data obtained, the most 1st component has a negative effect for all pairs. At the same time the 2nd component has a negative effect for the pairs: USD/RUB and ENR/RUB. For other couples, the effect of this companion is positive.

### Consider the data in the plane of the principal components PC1 and PC2.
```{r}
df_log$Year <- as.character(df$Date)
for (i in 1:nrow(df_log)) {
  df_log[i,5] = substr(df_log[i,5], 1, 4)
}
autoplot(pc, data = df_log, colour = 'Year')
```

### Construct a graph for a portfolio of currencies using only 2 principal components: PC1 and PC2.
```{r}
reconst <- t(t(pc$x[, 1:2] %*% t(pc$rotation[, 1:2]))
             * pc$scale + pc$center)
df_log <- df_log[,-5] 
diff <- df_log - reconst
#Stationarity test
adf.test(df_log$`USD/RUB`)
```

p.value above the acceptable level (5%). Consequently, the stationarity of the difference between the actual and calculated values is rejected. The strategy can be built on the basis of inertia in the behavior of actual and calculated values.

### Lets display the difference between the actual and the restored values.
### As example lets's use pair USD/RUB.
```{r}
plot(diff$`USD/RUB`, type = 'l', col = 'red')
```


### Let's imagine simulated data and actual data on the chart on the example of the USD/RUB pair
```{r}
reconst <- data.frame(reconst)
plot(reconst$USD.RUB, type = 'l', col = 'red')
lines(df_log$`USD/RUB`, type = 'l')
```

Analyzing the chart, it can be noted that a lot of time passes from the moment of one intersection to the next, which indicates the inertia of the course. This property will be used when building the strategy below.


## Use principal component method for strategy with basket of currency pairs.

### To develop the strategy, we will use the period to calculate 30 days and the requirement for a share of the explained variance of 90%.
```{r}
t <- 30  # The estimated window
D <- 0.9 # Required proportion of variance explained

N <- nrow(df) # Number of observations
C <- ncol(df_log) # Number of variables
```

### Let's create a matrix to fill with data.
```{r}
strawberry <- matrix(0, nrow = N, ncol = C) 
vanilla <- matrix(0, nrow = N, ncol = C) 
Npc <- rep(NA, N) 
```

### Strategy algorithm.
If the calculated value is higher than the actual, buy. If back, sell.
```{r}
for (i in (t+1):N) {
  temp <- df_log[(i-t):(i-1), 1:4] 
  pc <- prcomp(temp, scale. = TRUE) 
  pc_var <- pc$sdev^2 
  prop_var <- cumsum(pc_var/sum(pc_var)) 
  n <- which(prop_var > D)[1] 
  if (n != 2) next 
  Npc[i] <- n 
  reconst <- last(t(t(pc$x[, 1:n] %*% t(pc$rotation[, 1:n])) * pc$scale + pc$center)) 
  signal <- sign(last(temp) - reconst)  
  strawberry[i, 1:4] <- as.numeric(((df_log[i, 1:4] - df_log[i-1, 1:4]) / df_log[i-1, 1:4]) *signal)
  vanilla[i, 1:4] <- as.numeric(((df_log[i, 1:4] - df_log[i-1, 1:4]) / df_log[i-1, 1:4]))
}
```

### Let us compare the result of the strategy using the main components with the naive strategy on the criterion of cumulative profitability.
```{r}
a <- strawberry[,1] # Задаем значение стратегии
b <- vanilla[,1] # Задаем значения пары
c <- data.frame(x = a, y = b) # Создаем отдельный блок даннвх 
c$xa <- 1 
c$yb <- 1
for(i in 2:nrow(c)) { # Расчитываем прирос
  c[i,3] = c[i-1,3] * (1 + c[i,1])
  c[i,4] = c[i-1,4] * (1 + c[i,2])
}
plot(c$xa, type = 'l', ylim = c(0.6,2.5))
lines(c$yb, col = 'red')
```

The use of the strategy shows the same result as the usual holding of a portfolio consisting of currency pairs.

## Let's test the strategy using the principal components method for one currency USD / RUB.
```{r}
df_usd <- data.frame('USD' =  df$`USD/RUB`)
df_usd$USD <- as.numeric(df_usd$USD)
```

### Let's look at the profitability of the portfolio components consisting of currency pairs.
```{r}
m_p <- data.frame(strawberry)
mean(m_p$X1) # USD/RUB
mean(m_p$X2) # EUR/RUB
mean(m_p$X3) # GBP/RUB
mean(m_p$X4) # CNY/RUB
```

The yield of the USD/RUB pair is the only positive yield in the strategy. Accordingly, it is expected that if you use only the dollar/ruble pair, and the main components of their portfolio of currency pairs.

```{r}
strawberry_usd <- matrix(0, nrow = N, ncol = C) 
vanilla_usd <- matrix(0, nrow = N, ncol = C) 
Npc_usd <- rep(NA, N) 
```

```{r}
for (i in (t+1):N) {
  temp <- df_log[(i-t):(i-1), 1:4]  
  pc <- prcomp(temp, scale. = TRUE) 
  pc_var <- pc$sdev^2 
  prop_var <- cumsum(pc_var/sum(pc_var)) 
  n <- which(prop_var > D)[1]
  Npc_usd[i] <- n 
  reconst <- last(t(t(pc$x[, 1:n] %*% t(pc$rotation[, 1:n])) * pc$scale + pc$center)) 
  signal <- sign(last(temp) - reconst)
  strawberry_usd[i,] <- as.numeric(((df_usd[i,] - df_usd[i-1,]) / df_usd[i-1,]) *signal)
  vanilla_usd[i,] <- as.numeric(((df_usd[i,] - df_usd[i-1,]) / df_usd[i-1,]))
}
```

```{r}
a <- strawberry_usd[,1] 
b <- vanilla_usd[,1] 
c <- data.frame(x = a, y = b) 
c$xa <- 1 
c$yb <- 1
for(i in 2:nrow(c)) { 
  c[i,3] = c[i-1,3] * (1 + c[i,1])
  c[i,4] = c[i-1,4] * (1 + c[i,2])
}
plot(c$xa, type = 'l', ylim = c(0.6,2.5))
lines(c$yb, col = 'red')
```

## Summary:
### The main components defined in the portfolio consisting of currency pairs are a good method to use in the strategy.


